On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-08-26 Revisiones/PNA_Responses to the reviewers.tex

\documentclass[12pt]{article}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
\usepackage{graphicx}
\usepackage{cite}
\usepackage{color}
\usepackage{bm}

% \geometry{landscape} % rotated page geometry

\title{Response to the reviewers of the manuscript ``On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme"}
\author{Miguel Molinos, Pedro Navas, Manuel Pastor\\
and Miguel Mart\'in-Stickle}
\date{September 1, 2020}

@@

%%% BEGIN DOCUMENT
\begin{document}

\maketitle
%\tableofcontents

We  are grateful to the reviewers for taking their time  to review our work. Their comments have helped us to improve the paper. Changes to the original manuscript are given in  color (\textcolor{red}{red} for corrections, \textcolor{blue}{blue} for modified texts)  in the revised version.  A detailed response to both reviewers is given below.

\section*{Response to Reviewer \#1}
{\it
The authors have presented a promissing approach for MPM to surpass the shock wave propagation in the traditional MPM. The paper was very well written and the results suggested that the proposed approach works.
}
\\
\\
The reviewer's positive commentaries are greatly appreciated.
\\
\textit{However, most of the applications presented in this manuscript are relatively simple and does not really demonstrate the capability of the proposed approach. I suggest the authors to extend their manuscript to predict more challenging problems and compare their results to the original MPM solutions.}\\

Authors agree with the proposal of the reviewer. In the first version, aiming to make a readable manuscript, several examples were not shown. In this revised version, in order to clarify several questions of both reviewers, and in analogy with the work of Hammerquist \textit{et al.}\cite{HAMMERQUIST2017724}, where fast dynamics were also studied, the classical example of two colliding disks (first proposed by Sulsky) has been discussed.

@@

The development of the MPM approach has been revised considering the proposal of the reviewer.

\item \textit{There is little discussion on some of the pitfalls of using max-ent, e.g. the need for a convex domain, the issue of determining Lagrange Multiplier solutions for points close to boundaries, i.e. the solutions process here at Eqns 37 and 38.}\\

Further discussion has been added.

Of course, the employment of the Max-Ent shape function is not free of difficulties. Some requirements such as the need for a convex domain, as well as those derived from its calculation (the determination of the Lagrange Multipliers, specially at the boundaries, or the obtainment of the minimizer of the logarithmic function), make the usage of Max-Ent shape functions a challenging tool. However, the results depicted in this manuscript will strengthen the motivation of its employment in order to mitigate typical MPM problems such as cell-crossing or stress instabilities.

@@

\item \textit{Does the use of max-ent solve the problem of imposing Dirichlet boundary conditions in the MPM?}\\

Unfortunately, Local \textit{max-ent} approximation technique only provides an efficient tool to transfer information from particles to nodes and vice versa to the MPM. The difficulties of imposing Dirichlet boundary conditions (DBC) in the MPM are a direct consequence of discretising the continuum with material points instead of nodes, where traditionally DBC are imposed. See for instance the work of \cite{Cortis_et_al_2017_IJNME}. However, any available technique employed to overcome this situation in the standard MPM can be translated to the \textit{max-ent} one thanks to the Kronecker-Delta property inherent to \textit{max-ent}.

@@

\item \textit{I am sceptical of the novelty of the proposed explicit scheme as there is no mention of the recent XPIC(m) approach of Hammerquist and Nairn or similar approaches used in the Computational Graphics community, e.g. Stomakhin et al. (2013) or the more recent Polypic.}\\

These interesting publications will be added to the state of art of the revised manuscript. In the humble opinion of the authors, the XPIC approach consists of the stabilisation of the solution through a novel transference procedure more than a time integration scheme. Nevertheless, the algorithm introduced in this research proposes the stabilisation within the adaptation of the predictor-corrector Newmark algorithm, traditionally employed within FEM schemes, for the MPM framework. Hence, both techniques can be employed simultaneously.

@@

\end{enumerate}

 \textit{The numerical examples are not impressive when the message of the paper is that this is good for fast dynamics problems. Compare to papers such as Ma et al. (2009) IJ Impact Engineering}\\

Contrary to the reviewer's opinion, the provided results are capable to highlight the strong capabilities of the MPM Newmark Predictor-Corrector scheme. It is important to remark the extension of this research: in order to isolate the strength of the proposed methodology, covering spatial and time discretization, only linear elastic problems are assessed. Although the results presented in \cite{MA2009272} are impressive, the proposed examples are out of the scope of the present research. In order to reinforce the benefits of the proposed methodology and in analogy with the work of Hammerquist \textit{et al} \cite{HAMMERQUIST2017724} (previously cited by the reviewer), the classical example of two colliding disks proposed by Sulsky has been discussed.

### 2020-08-27 LME-MPM.tex
Descripción del caso de los dos discos

% Tikz
\usepackage{stanli}
\usepackage[ugly]{units}
% \usetikzlibrary{decorations}
% \usetikzlibrary{arrows}
\usetikzlibrary{plotmarks}

@@

This section is devoted to test the ability of both predictor-corrector
time integration scheme and the Local \textit{Max-Ent} approximants to
overcome spurious oscillations due to the grid crossing and high
frequency loads under the context of \acrshort{mpm}. Two different test have
been adopted for this purpose: the benchmark proposed by Dyka \& Ingel
(1995)\cite{Dyka1995} and the test proposed in the PhD thesis of
Andersen (2009)\cite{thesis_Andersen_2009}. On the one hand, the accuracy of the \acrfull{npc}
scheme is compared to the standard \acrfull{fe} in order to assess the performance of the time integration schemes. On the other hand,  \acrshort{lme} solutions
are compared with those provided by \acrfull{ugimp} and Q4 shape
functions. To avoid some mesh-dependent issues, in all
calculations a regular background mesh was set. All simulations were
performed with in-house software.

@@

The following test was proposed to validate the ability of the proposed
interpolation technique to deal with grid crossing instabilities. It
consists of the simulation of a square block of soil incrementally
loaded by a body force. Details of the problem are sketched in figure~\ref{fig:block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.7\hsize}{!}{
    \input{Figures/Block.tex}}
  \caption{Geometrical description of a soil block }.
  \label{fig:block}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This test was previously proposed by Andersen (2009)\cite{thesis_Andersen_2009}. The
elastic parameters considered for this test are: 
\begin{itemize} 
\item  Initial density : $6\cdot 10^3\ kg/m^3$
\item  Poisson ratio : $0$
\item  Elastic modulus : $5\ MPa$
\end{itemize}
The gravity force is applied as an external force according to the
equations \eqref{eq:particle_body_forces} and
\eqref{eq:nodal_external_forces}. Using a total time period of T (20
seconds) to apply the gravity, it is increased from 0 to 9.81$m/s$
with a sinusoidal function until T/2 and then maintained constant until T
in order to reach the equilibrium state:
\begin{equation}
  \label{eq:gravity-load-block}
 \mathbf{g}(t) = \left\{
    \begin{array}{ll}
      0.5 \mathbf{g} (\sin(\frac{2t \pi}{T} - \frac{\pi}{2})+1)  & \mbox{if } t \leq T/2 \\
      \mathbf{g} & \mbox{if } t > T/2
    \end{array}
  \right.
\end{equation}
In order to get a stable solution, time step was conducted by a
Courant number of 0.1. On the other hand, the explicit
predictor-corrector scheme is here employed looking forward to getting better results. For the
initial spatial discretisation four particles per cell
($\Delta x = 2\ m$) were adopted. The initial layout of particles inside of the
cell changes according to the approximation technique adopted. For the
bi-linear shape functions and the \acrshort{lme} approximants, the initial
position corresponds to the location of the gauss-points in a standard
quadratic finite element. For the \acrshort{ugimp} shape function the initial
position of each particle is located in the center of each voxel, due
to the fact that in the initial situation, the voxel domain should not
overlap each others.

@@

>>>>

% Tikz
% \usepackage{tikz}
%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
%setlength\PreviewBorder{5pt}
\usepackage{stanli}
\usepackage[ugly]{units}
% \usetikzlibrary{decorations}
% \usetikzlibrary{arrows}
\usetikzlibrary{plotmarks}

@@

This section is devoted to test the ability of both predictor-corrector
time integration scheme and the Local \textit{Max-Ent} approximants to
overcome spurious oscillations due to the grid crossing and high
frequency loads under the context of \acrshort{mpm}. \MODIFIED{Three different test have
been adopted for this purpose: the benchmark proposed by Dyka \& Ingel
(1995)\cite{Dyka1995}, the test proposed in the PhD thesis of
Andersen (2009)\cite{thesis_Andersen_2009} and the classical example of two colliding discs provided by Sulsky {\it et al.} (1994) \cite{Sulsky1994}. The first example is devoted to test the accuracy of the \acrfull{npc} scheme is compared to the standard \acrfull{fe} in order to assess the performance of the time integration schemes. In the second example,  \acrshort{lme} solutions
are compared with those provided by \acrfull{ugimp} and Q4 shape
functions. Finally, in the third example the present approach is tested under a scenario where shocks and grid crossing occurs.} To avoid some mesh-dependent issues, in all
calculations a regular background mesh was set. All simulations were
performed with in-house software.

@@

The following test was proposed to validate the ability of the proposed
interpolation technique to deal with grid crossing instabilities. It
consists of the simulation of a square block of soil incrementally
loaded by a body force. Details of the problem are sketched in figure~\ref{fig:block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.7\hsize}{!}{
    \input{Figures/Block.tex}}
  \caption{Geometrical description of a soil block }.
  \label{fig:block}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This test was previously proposed by Andersen (2009)\cite{thesis_Andersen_2009}. The
elastic parameters considered for this test are: 
\begin{itemize} 
\item  Initial density : $6\cdot 10^3\ kg/m^3$
\item  Poisson ratio : $0$
\item  Elastic modulus : $5\ MPa$
\end{itemize}
The gravity force is applied as an external force. Using a total time period of T (20
seconds) to apply the gravity, it is increased from 0 to 9.81$m/s$
with a sinusoidal function until T/2 and then maintained constant until T
in order to reach the equilibrium state:
\begin{equation}
  \label{eq:gravity-load-block}
 \mathbf{g}(t) = \left\{
    \begin{array}{ll}
      0.5 \mathbf{g} (\sin(\frac{2t \pi}{T} - \frac{\pi}{2})+1)  & \mbox{if } t \leq T/2 \\
      \mathbf{g} & \mbox{if } t > T/2
    \end{array}
  \right.
\end{equation}
In order to get a stable solution, time step was conducted by a
Courant number of 0.1. On the other hand, the explicit
predictor-corrector scheme is here employed looking forward to getting better results. For the
initial spatial discretisation four particles per cell
($\Delta x = 2\ m$) were adopted. The initial layout of particles inside of the
cell changes according to the approximation technique adopted. For the
bi-linear shape functions and the \acrshort{lme} approximants, the initial
position corresponds to the location of the gauss-points in a standard
quadratic finite element. For the \acrshort{ugimp} shape function the initial
position of each particle is located in the center of each voxel, due
to the fact that in the initial situation, the voxel domain should not
overlap each others.

@@

\subsection{Impact of two elastic discs}
\label{sec:Impact-two-elastic-discs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\sidecaption
 \centering
 \resizebox{0.7\hsize}{!}{
   \input{Figures/Sulsky_discs.tex}}
 \caption{Geometrical description of the two colliding discs problem.}
 \label{fig:Sulsky_discs}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
